Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-09-30 paper.tex

\usepackage{physics}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage[colorlinks=true, linkcolor=black, urlcolor=blue, citecolor=black, anchorcolor=black]{hyperref}

@@

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a$ and $\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
While most hyper-Raman experiments are performed using degenerate input pulses ($\omega_a = \omega_b$), it is possible to perform the technique with non-degenerate input pulses ($\omega_a \neq \omega_b$). \cite{Denisov1986, Kozich2007}
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through a two-photon inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies (sometimes under the moniker of SIVE).\cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Wang2021, Bonn2024, McDonnell2024}

@@

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares them with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require macroscopic non-centrosymmetry under the electric dipole approximation.\cite{RN480}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to four-wave mixing (FWM) cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also weak and highly susceptible to FWM cascades.\cite{Taran1977, Berger1978, Bjarnason1980, Cho1997, Cho1998}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave and six-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling).\cite{RN445, RN335} 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as two-color experiments (with $\omega_2=\omega_3$);\cite{Cho2001} vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate ev coupling in bulk systems.

Nonlinear spectroscopic probes of electronic-vibrational coupling are not limited to the Raman and hyper-Raman based techniques diagrammed in \autoref{fig:comparisonwmel}.
Independently developed by the Fleming and Khalil groups, respectively, 2D Electronic-Vibrational (2D-EV) and 2D Vibrational-Electronic (2D-VE) spectroscopy are based upon pump-probe and absorption type pathways which also investigate e-v coupling.\cite{Oliver2014, Courtney2015, Courtney2015_1}
The Rao group has also developed a $\chi^{(4)}$ analogue of 2D-EV, 2D-EV sum frequency generation (2D-EVSFG), to probe interfacial e-v coupling. \cite{Deng2021}  
These methods have been used to understand vibronic coupling in a variety of molecular systems and, albeit different selection rules, serve as 2D pump-probe type analogues to the hyper-Raman-based spectroscopies discussed here.\cite{Courtney2015, Gaynor2018, Arsenault2020, HuangFu2023}

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
Coherent spectroscopies are often described using the density matrix formalism, where an n$^{th}$ order density matrix $\rho^{(n)}$ is propagated using the Liouville-von Neumann equation $via$
\begin{equation}\label{LVNE}
	\dot{\rho}^{(n)} = \frac{1}{i \hbar} [H + V(t), \rho^{(n-1)}] + \dot{\rho}^{(n)}_R,
\end{equation}
where $H$ is the (static) system Hamiltonian, $V(t)$ is the field-matter interaction and $\rho_R$ is the relaxation density matrix. \cite{RN455}
In systems where interactions with surroundings are relevant, $\dot{\rho}_R$ becomes a complex function that can be modified to include phenomena such as system-bath interactions, solvent-solute coupling, among others. \cite{Sung2001}
In the limit where input field pulsewidths are much longer than electronic dephasing times and the individual molecule limit is approached, it becomes acceptable to write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$, i.e. exponential relaxation, where $\Gamma$ is the system's dephasing rate. \cite{RN455, Sue1986, Sung2001}
While exponential relaxation is usually an inappropriate description of most physical systems involving solvent-solute interactions,\cite{Sue1986, Yan1988, Li1994, Myers1997} our goal is to introduce simple, quantitative models to understand HDFG response.
Thus, we write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$.
The exponential relaxation limit has proven useful in investigating the properties of 2D-EV, 2D-VE, and doubly resonant SFG using pulses much longer than electronic dephasing times. \cite{Raschke2002, Gaynor2017}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{RN287}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
		The virtual states are arbitrarily placed and fluctuate in energy dependent upon the energy of the input fields.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned. \cite{RN287, McDonnell2024}
Note that in the case where $\omega_1$ interacts first but $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

@@

Another potential interferant in HDFG spectroscopy is direct and sequential cascades of second-order vDFG and vSFG processes.\cite{RN297, RN301}
While unimportant in achiral isotropic systems,\cite{Belkin2000} second-order cascades may become important in media where SFG and DFG are allowed, e.g. chiral media, interfaces or non-centrosymmetric media. 
Nevertheless, in the discussion that follows for the rest of this manuscript, we will assume negligible second-order cascades.

@@

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,\tilde{v}'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | \tilde{v}' \rangle
		 \langle \tilde{v}' | 0 \rangle 
		 \frac{1}{\Delta_{e\tilde{v}', gv}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | \tilde{v}' \rangle \mel{\tilde{v}'}{Q_z}{0} \frac{1}{\Delta_{e\tilde{v}', gv}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{\tilde{v}'} 
			\langle \tilde{v}' | 0 \rangle 
			\frac{1}{\Delta_{e\tilde{v}', gv}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
Note that it is common in the hyper-Raman literature to write $\Delta_{e\tilde{v}', gv} \Delta_{mn, gv} = \Delta_{e\tilde{v}', gv}$.
Here, $\langle \tilde{a} | b \rangle$ and $\mel{\tilde{a}}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively.
Duschinsky rotation effects are ignored in this discussion. \cite{Duschinsky1937}
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

To understand the vibronic structure of electronic resonance, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode, $\ket{g,1}$, so that hyper-Raman scattering is between ground state and $\ket{g,1}$. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{\tilde{v}'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | \tilde{v}' \rangle
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',gv}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar} \sum_{\tilde{v}'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | \tilde{v}' \rangle \mel{\tilde{v}'}{Q}{0} 
				\frac{1}{\Delta_{e\tilde{v}', gv}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{\tilde{v}'} 
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',gv}}
			\end{split}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $ is the electronic two photon transition moment between $|a)$ and $|b)$.\cite{McClain1977}

@@

@@

\autoref{ABterms_DR} shows important selection rules for vibronic transitions in HDFG.
The $A$ term is allowed whenever an electronic transition has a one- and two- photon transition moment.
For centrosymmetric molecules, an electronic transition cannot be both one- and two- photon allowed by parity, so the $A$ term will vanish.
This makes HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects.\cite{Milojevich2011, Olson2018}
For an allowed electronic transition, the brightness of a vibronic resonance is controlled by overlap of ground and excited state vibrational wavefunctions (Franck-Condon factors).
The $A$ term relies on excited state perturbations to the vibrational wavefunctions; if the excited electronic state is minimally offset ($\Delta \rightarrow 0$), $\langle 1 | \tilde{v}' \rangle \langle \tilde{v}' | 0 \rangle = 0$ and $A$ vanishes, as discussed in the supporting information.
Some $B$ terms do survive in the same limit, however; for instance, $B_1$ survives for $\tilde{v}'=1$ and $B_2$ survives for $\tilde{v}'=0$.
In the absence of an $A$ term, the $B$ terms will dominate.
The $B_1$ and $B_2$ terms involve different types of vibronic coupling pathways.
In $B_1$, the Herzberg-Teller coupling is associated with the two-photon absorption process, whereas in $B_2$, the Herzberg-Teller coupling is associated with one-photon emission.\footnote{Franck-Condon factors and Herzberg-Teller integrals are most easily evaluated in the position basis. A sample calculation is provided in the supporting information.}

\autoref{fig:doubres_spec}b shows the simulated $\omega_2+\omega_3$ spectrum for this two-well system.  The spectrum selectively probes vibronic pathways from $\ket{g,0}$ to $\ket{g,1}$.
In the resonant region, the $A$ and $B$ terms contain character from all three possible transitions between $\ket{g,0}$, $\ket{e,\tilde{v}'}$ and $\ket{g,1}$. 
Significant contributions from $\ket{e,0}$ and $\ket{e,1}$ in the $A$ term spectrum correspond well with expectations given the relative wavefunction overlap in \autoref{fig:doubres_spec}a, tabulated in \autoref{T:contributions}. 
Similarly, the $B_1$ term is dominated by coupling with the $\ket{e,1}$ state and $B_2$ is dominated by coupling with the $\ket{e,0}$ state (\autoref{T:contributions}). 
Note that contributions involving $\ket{e,2}$ are suppressed relative to the $\ket{e,0}$ and $\ket{e,1}$ contributions in the $A$ and $B$ terms.
This is expected, considering the large change in quantum number when involving $\ket{e,2}$ and the offset of the potential wells in \autoref{Hamiltonian}.

@@

Since $\Delta$ can take on a variety of values given a specific system, in the supporting information we demonstrate the impact of $\Delta$ on the HDFG response using previously reported values for some Raman active vibrations in \textit{trans}-hexatriene and UiO-68 metal organic frameworks (Figure S2). \cite{Myers1982, Brennan2024}
Clearly, specific $\Delta$ and $\omega_{g1, g0}$ values have a significant impact on the HDFG spectral profile.
Indeed, for some values of $\Delta$, it appears that $B$ term resonances dominate over $A$ term resonances, which can allow for significant non-Condon behavior in pre-resonant HDFG response.

In the post-resonance region, the amplitude takes on an unexpected lineshape.
This arises due to interference between the real and imaginary parts of $\gamma$ (\autoref{fig:doubres_spec}c, d).
The abnormal lineshape arises due to the sign of the Franck-Condon and Herzberg-Teller integrals, as well as the molecular parameters that drive the output of the vibronic transitions.
Such effects have been seen elsewhere in spontaneous, surface enhanced hyper-Raman studies. \cite{Milojevich2011}

The Franck-Condon and Herzberg-Teller integrals can manipulate interference in the total output spectrum as a function of offset (Figure S2).
When the offset is inverted about the origin ($\Delta \rightarrow -\Delta$),

However, the main source of interference in the HDFG spectra arises from the sign of the molecular parameters. 
In the HDFG spectrum presented in \autoref{fig:doubres_spec}, the signs of all molecular parameters are taken to be positive. 
Inverting the sign of one parameter eliminates the interference and restores the expected, purely Lorentzian lineshape.

>>>>

\usepackage{physics}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
%\usepackage{times}
\usepackage{etoolbox}
\usepackage[colorlinks=true, linkcolor=black, urlcolor=blue, citecolor=black, anchorcolor=black]{hyperref}

\usepackage{lineno}
\linenumbers

@@

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a$ and $\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
While most hyper-Raman experiments are performed using degenerate input pulses ($\omega_a = \omega_b$), it is possible to perform the technique with non-degenerate input pulses ($\omega_a \neq \omega_b$). \cite{Denisov1986, Kozich2007}
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through a two-photon inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies (sometimes under the moniker of SIVE).\cite{Zilian1994, RN350, RN351, RN352, RN353, Chen1998, RN362, RN418, Wang2021, Bonn2024, McDonnell2024}

@@

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares them with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require macroscopic non-centrosymmetry under the electric dipole approximation.\cite{RN480}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to four-wave mixing (FWM) cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also weak and highly susceptible to FWM cascades.\cite{Taran1977, Berger1978, Bjarnason1980, Cho1997, Cho1998}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave and six-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling).\cite{RN445, RN335} 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as two-color experiments (with $\omega_2=\omega_3$);\cite{Cho2001} vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements to investigate ev coupling in bulk systems.

Nonlinear spectroscopic probes of electronic-vibrational coupling are not limited to the Raman and hyper-Raman based techniques diagrammed in \autoref{fig:comparisonwmel}.
Independently developed by the Fleming and Khalil groups, respectively, 2D Electronic-Vibrational (2D-EV) and 2D Vibrational-Electronic (2D-VE) spectroscopy are based upon pump-probe and absorption type pathways which also investigate e-v coupling.\cite{Oliver2014, Courtney2015, Courtney2015_1}
The Rao group has also developed a $\chi^{(4)}$ analogue of 2D-EV, 2D-EV sum frequency generation (2D-EVSFG), to probe interfacial e-v coupling. \cite{Deng2021}  
These methods have been used to understand vibronic coupling in a variety of molecular systems and, albeit different selection rules, serve as 2D pump-probe type analogues to the infrared-hyper-Raman spectroscopies discussed here.

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
Coherent spectroscopies are often described using the density matrix formalism, where an n$^{\text{th}}$ order density matrix $\rho^{(n)}$ is propagated using the Liouville-von Neumann equation $via$
\begin{equation}\label{LVNE}
	\dot{\rho}^{(n)} = \frac{1}{i \hbar} [H + V(t), \rho^{(n-1)}] + \dot{\rho}^{(n)}_R,
\end{equation}
where $H$ is the (static) system Hamiltonian, $V(t)$ is the field-matter interaction and $\rho_R$ is the relaxation density matrix. \cite{RN455}
In systems where interactions with surroundings are relevant, $\dot{\rho}_R$ becomes a complex function that can be modified to include phenomena such as system-bath interactions, solvent-solute coupling, among others. \cite{Sung2001}
In the limit where input field pulsewidths are much longer than electronic dephasing times and the individual molecule limit is approached, it becomes acceptable to write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$, i.e. exponential relaxation, where $\Gamma$ is the system's dephasing rate. \cite{RN455, Sue1986, Sung2001}
While exponential relaxation is usually an inappropriate description of most physical systems involving solvent-solute interactions,\cite{Sue1986, Yan1988, Li1994, Myers1997} our goal is to introduce simple, quantitative models to understand HDFG response.
Thus, we write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$.
The exponential relaxation limit has proven useful in investigating the properties of 2D-EV, 2D-VE, and doubly resonant SFG using pulses much longer than electronic dephasing times. \cite{Raschke2002, Gaynor2017}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{Cho2008}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
		The virtual states are arbitrarily placed and fluctuate in energy dependent upon the energy of the input fields.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned. \cite{McDonnell2024}
Note that in the case where $\omega_1$ interacts first but $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

@@

Other potential interferants in HDFG spectroscopy are direct and sequential cascades of second-order vDFG and vSFG processes.\cite{RN297, Cho2000_Cascade, RN301}
While unimportant in achiral isotropic systems,\cite{Belkin2000} second-order cascades may become important in media where SFG and DFG are allowed, e.g. chiral media, interfaces or non-centrosymmetric media. 
Nevertheless, in the discussion that follows for the rest of this manuscript, we will assume negligible second-order cascades.

@@

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,\tilde{v}'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | \tilde{v}' \rangle
		 \langle \tilde{v}' | 0 \rangle 
		 \frac{1}{\Delta_{e\tilde{v}', gv}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | \tilde{v}' \rangle \mel{\tilde{v}'}{Q_z}{0} \frac{1}{\Delta_{e\tilde{v}', gv}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,\tilde{v}',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{\tilde{v}'} 
			\langle \tilde{v}' | 0 \rangle 
			\frac{1}{\Delta_{e\tilde{v}', gv}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
It is assumed that all derivatives above are evaluated at $Q_z = 0$.
Note that it is common in the hyper-Raman literature to write $\Delta_{e\tilde{v}', gv} \Delta_{mn, gv} = \Delta_{e\tilde{v}', gv}$.
Here, $\langle \tilde{a} | b \rangle$ and $\mel{\tilde{a}}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively.
Duschinsky rotation effects are ignored in this discussion. \cite{Duschinsky1937}
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

To understand the vibronic structure of an excited state $|e)$, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode, $\ket{g,1}$, so that hyper-Raman scattering is between ground state and $\ket{g,1}$. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
				A_{ijk} = \frac{1}{\hbar}\sum_{\tilde{v}'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | \tilde{v}' \rangle
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',gv}}
		\end{equation}
		\begin{equation}
				B_{1_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | \tilde{v}' \rangle \mel{\tilde{v}'}{Q}{0} 
				\frac{1}{\Delta_{e\tilde{v}', gv}}
		\end{equation}
		\begin{equation}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{\tilde{v}'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{\tilde{v}'} 
				\langle \tilde{v}' | 0 \rangle 
				\frac{1}{\Delta_{e\tilde{v}',gv}}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $ is the electronic two photon transition moment between $|a)$ and $|b)$.

@@

It is important to realize that these expansions are taken about the equilibrium point of the ground state surface, i.e., $Q = 0$.
Some recent studies investigating interference between Franck-Condon and Herzberg-Teller terms in absorption and emission spectra have found it useful to Taylor expand about the midpoint of the equilibria points of the ground and excited surface.\cite{Kundu2022}
If such a definition was used here, $B$ would possess non-negligible Franck-Condon character.
As such, it is more useful to assign Herzberg-Teller character to the $B$ term due to the presence of $\frac {\partial M^{ge}}{\partial Q}$ and $\frac {\partial \Lambda^{ge}}{\partial Q} $ terms.
Nevertheless, since the molecular response must be invariant to the center point of the Taylor expansion, expanding about $Q = 0$ or the midpoint should provide similar results; if the dipole surface has a kink or node about $Q = 0$, then the midpoint would be a better choice for understanding Herzberg-Teller contributions to the spectra.

@@

\autoref{ABterms_DR} shows important selection rules for vibronic transitions in HDFG.
The $A$ term is allowed whenever an electronic transition has a one- and two- photon transition moment.
For centrosymmetric molecules, an electronic transition cannot be both one- and two- photon allowed by parity, so the $A$ term will vanish.
This makes HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects.\cite{Milojevich2011, Olson2018}
For an allowed electronic transition, the brightness of a vibronic resonance is controlled by overlap of ground and excited state vibrational wavefunctions (Franck-Condon factors).
The $A$ term relies on excited state perturbations to the vibrational wavefunctions; if the excited electronic state is minimally offset ($\Delta \rightarrow 0$), $\langle 1 | \tilde{v}' \rangle \langle \tilde{v}' | 0 \rangle = 0$ and $A$ vanishes, as discussed in the supporting information.
Some $B$ terms do survive in the same limit, however; for instance, $B_1$ survives for $\tilde{v}'=1$ and $B_2$ survives for $\tilde{v}'=0$.
In the absence of an $A$ term, the $B$ terms will significantly dominate.
The $B_1$ and $B_2$ terms involve different types of vibronic coupling pathways.
In $B_1$, the Herzberg-Teller coupling is associated with the two-photon absorption process, whereas in $B_2$, the Herzberg-Teller coupling is associated with one-photon emission.\footnote{Franck-Condon factors and Herzberg-Teller integrals are most easily evaluated in the position basis. A sample calculation is provided in the supporting information.}
For small $\Delta$ values, it is clear by comparing the relative Franck-Condon and Herzberg-Teller integral products that $B$ term resonances can dominate the spectrum (Figure S2).
Similar ideas have been observed in resonance hyper-Raman spectra, which makes HDFG a sensitive probe of non-Condon effects for those modes with small $\Delta$ values.

\autoref{fig:doubres_spec}b shows the simulated $\omega_2+\omega_3$ spectrum for this two-well system.  The spectrum selectively probes vibronic pathways from $\ket{g,0}$ to $\ket{g,1}$.
In the resonant region, the $A$ and $B$ terms contain character from all three possible transitions between $\ket{g,0}$, $\ket{e,\tilde{v}'}$ and $\ket{g,1}$. 
Significant contributions from $\ket{e,0}$ and $\ket{e,1}$ in the $A$ term spectrum correspond well with expectations given the relative wavefunction overlap in \autoref{fig:doubres_spec}a, tabulated in \autoref{T:contributions}. 
The overlaps are plotted as function of offset in Figure S2.
Similarly, the $B_1$ term is dominated by coupling with the $\ket{e,1}$ state and $B_2$ is dominated by coupling with the $\ket{e,0}$ state (\autoref{T:contributions}). 
Note that contributions involving $\ket{e,2}$ are suppressed relative to the $\ket{e,0}$ and $\ket{e,1}$ contributions in the $A$ and $B$ terms.
This is expected, considering the large change in quantum number when involving $\ket{e,2}$ and the offset of the potential wells in \autoref{Hamiltonian}.

@@

Since $\Delta$ can take on a variety of values given a specific system, in the supporting information we demonstrate the impact of $\Delta$ on the HDFG response using previously reported values (Figure S3). \cite{Myers1982, Brennan2024}
Clearly, $\Delta$, $\omega_{g1, g0}$ and $\Gamma$ have a significant impact on the HDFG spectral profile.
Indeed, for some values of $\Delta$, it appears that $B$ term resonances dominate over $A$ term resonances (see also Figure S2), which can allow for significant non-Condon behavior in pre-resonant HDFG response.
